Final Project Part 3 Report

### 2024-12-07 final_paper/final_report.Rmd
inital

Housing affordability has become a pressing issue in many urban cities, including Toronto, where rising house prices have placed both homeowners and tenants out of reach for many years. This paper is conducted to investigating several key house indicators to the influence of house price. This is essential for policymakers , real estate professionals and potential homeowners seeking to navigate this challenging market.

@@

We conducted literature reviews to validate the selection of our predictors based on prior studies. Housing supply is consistently shown to be a critical factor influencing house prices. Research indicates that higher sales activity (houses sold) is often associated with increased demand, which drives prices upward. Conversely, an oversupply of unsold houses tends to push prices downward. These relationships are commonly examined using hedonic and linear regression models, where transactional data serve as key predictors for estimating price changes effectively (Geerts et al., 2023).

The number of newly constructed and completed houses influences house prices by affecting market supply. An increase in construction tends to stabilize prices by meeting demand, while lower construction rates can contribute to supply constraints and higher prices. "Global House Prices: Trends and Cycles" highlights the influence of new constructions and completions on house prices (Gnan, 2021). Regression analyses reveal that increases in supply through construction stabilize prices, while supply shortages contribute to price hikes

In the context of CPI, the CEPR report discusses their inclusion in econometric models to study the effects of borrowing costs on housing affordability and prices (Muellbauer & Murphy, 2021). It finds that higher CPI indicates higher inflation, reflected in higher mortgage rates and reducing both housing demand and price growth, a relationship frequently captured through regression techniques

@@

For this analysis, the focus will be on identifying and validating the linear relationship between the predictors and the response variable. We will employ exploratory data analysis (EDA) and Variance Inflation Factor (VIF), evaluate model performance using metrics such as AIC, BIC, $R^2$, and adjusted $R^2$. Through these statistical methods, we aim to establish that a linear regression relationship exists between house prices and the predictors: CPI, unabsorbed homes, and completed construction homes.

>>>>

Housing affordability has become a pressing issue in many urban cities, including Toronto, where rising house prices have placed both homeowners and tenants out of reach for many years. This paper is conducted to investigating several key house indicators to the influence of house price. This is essential for policymakers, real estate professionals and potential homeowners seeking to navigate this challenging market.

@@

We conducted literature reviews to validate the selection of our predictors based on prior studies. Housing supply is consistently shown to be a critical factor influencing house prices. Research indicates that higher sales activity (houses sold) is often associated with increased demand, which drives prices upward. Conversely, an oversupply of unsold houses tends to push prices downward. These relationships are commonly examined using hedonic and linear regression models, where transactional data serve as key predictors for estimating price changes effectively (Geerts et al., 2023).
The number of newly constructed and completed houses influences house prices by affecting market supply. An increase in construction tends to stabilize prices by meeting demand, while lower construction rates can contribute to supply constraints and higher prices. "Global House Prices: Trends and Cycles" highlights the influence of new constructions and completions on house prices (Gnan, 2021). Regression analyses reveal that increases in supply through construction stabilize prices, while supply shortages contribute to price hikes.

In the context of CPI, the CEPR report discusses their inclusion in econometric models to study the effects of borrowing costs on housing affordability and prices (Muellbauer & Murphy, 2021). It finds that higher CPI indicates higher inflation, reflected in higher mortgage rates and reducing both housing demand and price growth, a relationship frequently captured through regression techniques.

@@

For this analysis, the focus will be on identifying and validating the linear relationship between the predictors and the response variable. We will employ exploratory data analysis (EDA) and Variance Inflation Factor (VIF), evaluate model performance using metrics such as AIC, BIC, $R^2$, adjusted $R^2$ and RMSE. Through these statistical methods, we aim to establish that a linear regression relationship exists between house prices and the predictors: CPI, unabsorbed homes, and completed construction homes.
